---
title: "Data 607 HW 9"
author: "John Perez"
date: "3/30/2019"
output: html_document
---


## Assignment Web APIs

Below, we will connect to the New York Times movie reviews API, search for movies containing the string "Star Wars", and create a dataframe.  

```{r, include=FALSE, warning=FALSE, message=FALSE}
library(stringr)
library(jsonlite)
library(dplyr)

```


Here I declare a function that takes a string keyword corresponding to the movie we want to search for and ouputs the URL that returns those results. 

The fromJSON() function is used to read the data, and a dataframe is created with the following columns:  

+ title
+ mpaa_rating
+ byline
+ summary_short
+ opening_date
+ url

```{r connect NYT}
nytquery <- function(keywords){
    queryelem <-  str_c(c("query=",URLencode(keywords)), collapse="")
    apiKey <- "UR7fDqc0nEUYw8XFye8qtSGOo5zTH8PR"
    base_url <- "http://api.nytimes.com/svc/movies/v2/reviews/search.json?api-key="
    url <- str_c(c(base_url, apiKey,"&", 
    queryelem), collapse = "")
    return(url)
}

query <- nytquery("Star Wars")
results <- fromJSON(query)

results_df <- results$results %>% subset(select=c(display_title,mpaa_rating,byline,summary_short,opening_date))

results_df$url <- results$results$link$url
results_df$opening_date <- as.Date(results_df$opening_date)
results_df <- results_df[order(results_df$opening_date),]

DT::datatable(results_df)
```

