---
title: "Data 607 HW 3"
author: "John Perez"
date: "2/16/2019"
output: html_document
---

## Question 3  
  
(a) Conform all elements to standard first_name last_name
```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
df<-data.frame(name = name, phone = phone, stringsAsFactors=FALSE)

for(i in 1:nrow(df)) {
    if(str_detect(df$name[i],",")){
        split <- strsplit(df$name[i], split = ", ")
        df$name[i]<-str_c(split[[1]][2], " ", split[[1]][1])
    }
}
df
```
(b) Construct a logical vector indicating whether a character has a title 
```{r}
title <- str_detect(df$name, "\\w{2,}\\.")
title
```
(c) Construct a logical vector indicating whether a character has a second name 
```{r}
second_name <- c()
for(i in 1:nrow(df)) {
    if(length(strsplit(df$name[i],split=" ")[[1]]) > 2){ #if greater than 2 words
        if(str_detect(df$name[i], "[A-Z]\\.")){ #if 1 letter abbrv. mark as true
            second_name[i]<-TRUE
        }else{
            second_name[i]<-FALSE
        }
    }else{
        second_name[i]<-FALSE
    }
}
second_name
```

## Question 4  
(a) This looks for one or more digits at the end of a string. ex: "1234" 
(b) This looks for any lowercase word between 1 and 4 characters. ex: "test"
(c) This looks for any string ending in ".txt". ex: "123abc.txt"
(d) This looks for any 2 digits followed by slash, followed by 2 digits, followed by slash, followed by 4 digits. ex: "11/23/1999"
(e) This looks for a "<" character followed by any characters followed by ">". It then matches any characters, then a "</" followed by the same result from the first group. ex: "\<anything\>sometext\</anything\>"

## Question 9

```{r}
text <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
message <- unlist(str_extract_all(text, "[[:upper:].]"))
message <- str_replace_all(paste(message, collapse=''),"\\."," ")
message
```

